\speculake: Hiding Malware in Speculative Execution

### 2018-06-26 paper/architecture.tex
minor fixes including spelling and phrasing requests from reviewers

We found that the processor can identify some operations that have no effect on
output registers and allows them to use the entire size provided by the ROB, as
determined in Section~\ref{sssec:ROB}. This includes idomatic no-ops and zero
idioms, which have no reliance on register values, for example the instruction
\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
instruction is a potential data hazard, it must use an entry in the PRF until it
reaches the correct stage in the execution pipeline. In this case, the number of
instructions could be limited by the PRF entries available.

>>>>

We found that the processor can identify some operations that have no effect on
output registers and allows them to use the entire size provided by the ROB, as
determined in Section~\ref{sssec:ROB}. This includes idomatic no-ops and zero
idioms, which have no reliance on register values, for example the instruction
\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
instruction is a potential data hazard, it must use an entry in the Physical 
Register File (PRF) until it reaches the correct stage in the execution pipeline. 
In this case, the number of instructions could be limited by the PRF
entries available.

### 2018-06-26 paper/discussion.tex
minor fixes including spelling and phrasing requests from reviewers

\texttt{Guest1 -> Guest1}---When both trigger and payload are run internal to a
single virtual machine the branch prediction is successfully trained in the
payload program, though at a notably reduced hit rate. This can be overcome by
using an increased number of iterations to establish confidence.

\texttt{Guest -> Host}---In a configuration where trigger runs on the guest
while the payload runs on the host we have found that \speculake also
successfully trains the branch predictor well enough retreive a signal. Again,
this signal is significantly reduced, implying the hypercall context swtich from
guest to host requires some overhead. We note that processor affinity in a guest
is not mapped to processor affinity in the host, as the host assigns the VM to
any available CPU, though eventually the hypervisor will assign the VM back to
the same core as the payload program allowing progress to continue.

\texttt{Host -> Guest}---When trigger runs on the host attempting to train the
branch predictor to influence a payload process in a guest we have found that
the payload receives no signal. This implies that the hypercall context switch
from the host to the guest is heavier and has a larger impact on branch history.
This matches with work done by the Project Zero team at Google which found that
hypervisors can be fingerprinted from within a VM using the branch
predictor~\cite{project_zero}.

\texttt{Guest1 -> Guest2}---Configurations where trigger and payload run on
co-resident virtual machines similarly show no signal for the same reasons that
host to guest \speculake fails. However, this has the added difficulty that
virtual machines must be run on the same core or on a pair of hyperthreaded
cores.

@@

\subsubsection{Benign Triggers}
We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
\speculake payloads to be triggered remotely, and discussed the minimum
requirements for identifying a benign trigger. However, we have not classified
typical system resources that might be used as benign triggers. This is a
promising direction for future investigation as various system libraries present
juicy targets for trigger design. For exmaple, if a sufficient series of jumps
is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
TCP Checksum library) an attacker could reasonably trigger a \speculake payload
remotely, from a spoofed IP address. This attack would rely on a payload
defeating KASLR, however, this has been demonstrated to be
feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.

>>>>

\texttt{Guest1 -> Guest1}---When both the trigger and payload are run internal to a
single virtual machine the branch prediction is successfully trained in the
payload program, though at a notably reduced hit rate. This can be overcome by
using an increased number of iterations to establish confidence.

\texttt{Guest -> Host}---In a configuration where the trigger runs on the guest
while the payload runs on the host we have found that \speculake also
successfully trains the branch predictor well enough retrieve a signal. Again,
this signal is significantly reduced, implying the hypercall context switch from
guest to host requires some overhead. We note that processor affinity in a guest
is not mapped to processor affinity in the host, as the host assigns the VM to
any available CPU, though eventually the hypervisor will assign the VM back to
the same core as the payload program allowing progress to continue.

\texttt{Host -> Guest}---When the trigger program executes on the host
attempting to train the
branch predictor to influence a payload process in a guest we have found that
the payload receives no signal. This implies that the hypercall context switch
from the host to the guest is heavier and has a larger impact on branch history.
This is also supported by the fact that hypervisors can be fingerprinted from 
within a VM using the branch predictor as demonstrated by the Project Zero 
team at Google~\cite{}.
~\cite{project_zero}.

\texttt{Guest1 -> Guest2}---Configurations where the trigger and payload run on
co-resident virtual machines similarly show no signal for the same reasons that
host to guest \speculake fails. However, this has the added difficulty that
virtual machines must be run on the same core or on a pair of hyperthreaded
cores.

@@

\subsubsection{Benign Triggers}
We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
\speculake payloads to be triggered remotely, and discussed the minimum
requirements for identifying a benign trigger. However, we have not classified
typical system resources that might be used as benign triggers. This is a
promising direction for future investigation as various system libraries present
juicy targets for trigger design. For example, if a sufficient series of jumps
is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
TCP Checksum library) an attacker could reasonably trigger a \speculake payload
remotely, from a spoofed IP address. This attack would rely on a payload
defeating KASLR, however, this has been demonstrated to be
feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.

### 2018-06-26 paper/related-work.tex
minor fixes including spelling and phrasing requests from reviewers

The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
composing bugs and features of a system to create arbitrary compuatation with
interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}.
This encompasses many traditional computer vulnerabilities including buffer
overflows~\cite{buffer_overflow}, format string
exploits~\cite{format_string_exploit}, and return oriented
programming~\cite{shacham2007geometry}. These exploits demonstrate performing
arbitrary computation in places that were not originally intended to act as
computational entities.

>>>>

The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
composing bugs and features of a system to create arbitrary computation with
interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}.
This encompasses many traditional computer vulnerabilities including buffer
overflows~\cite{buffer_overflow}, format string
exploits~\cite{format_string_exploit}, and return oriented
programming~\cite{shacham2007geometry}. These exploits demonstrate performing
arbitrary computation in places that were not originally intended to act as
computational entities.
